---
title: "The_final_code"
author: "Anders Havbro, Morten Gade, Sigrid Nielsen and Gustav Helms"
date: "2/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Setup
library(pacman)
pacman::p_load("tidyverse")

#Loading the data
df <- read.csv("sleepstudy.csv", header = T)
```

## Task 1
### 1.a: Get the data from one participant, e.g. using subset(). Make a linear regression for reaction time as a function of days of sleep deprivation, e.g. using lm(). Report the F-statistics.

### 1.b: How many degrees of freedom does the relevant F-distribution have?

### 1.c: At which F-value does a regression with this distribution become statistically significant (p<0.05)?

### 1.d: Make a plot of the F-distribution.

```{r}

```

## Task 2. For all participant in the experiment
### 2.a: Find the coefficients (slope and intercept) for the regression for reaction time as a function of days of sleep deprivation (a hint for the solution: use group_by() in tidyverse or this function here: https://stat.ethz.ch/R-manual/R-devel/library/nlme/html/lmList.html,hint2: pool=FALSE)

### 2.b: Combine both scatter plot and regression line in the same figure. You may also include all participants in one plot.

### 2.c: Collect and report the inferential statistics for each participant in a table using t-statistics, including t-value, df and p-value.

### 2.d: How many individual participants display a statistically significant effect of sleep deprivation (p-values uncorrected for mulitple comparisons)?

```{r}

```

## Task 3: Across participants: 
### 3.a: Use the slopes you found for each participant in exercise 2 as a new dataset. Test the hypothesis that the slopes are larger than zero against the null-hypothesis that the slopes are zero (i.e. no differences in response time exist as a function of time).

### 3.b: Justify your use of test statistics.

### 3.c: Report inferential statistics.

### 3.d: Make a plot with the mean reaction time and standard errror bars for each day across participants and plot the averaged regression line in the same figure.

```{r}

```

