---
title: "Portofolio 4"
author: "Gustav Helmet, Morten Street, Anders Wheelman & Sigrid Snapfield"
date: "2/26/2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Setup
library(pacman)
pacman::p_load("tidyverse", "lme4", "reshape2", "pracma","Hmisc")

#Importing data
alignment <- read.delim("alignment.txt", header = F, sep = "")

#Importing the designmatrix
fmrides<-as.matrix(read.csv("aud_fmri_design.csv", header=FALSE))
##making it a time-series
fmrides2<-ts(fmrides)

```


## 3. Realignment parameters
SPM produces a file with the realignment parameters, i.e. the calcultated participant movement and rotation per time point (called “rp. . . txt”, saved together with the functional data).

Import the realignment parameters into R.

### 3.a. Make a lineplot of the realignment parameters in R.
```{r}
#Transforming the dataframe into a melted format
alignment <- melt(alignment)

#Making a new time variable for the x-axis
alignment$Time <- c(1:400)

#Plotting 
ggplot(alignment, aes(Time, value, colour = variable))+
  geom_line()+
  ggtitle("Realignment parameters")
```


### 3.b. How far has the participant moved for each dimension during the experiment (Hint: use “apply()” to run a function across columns)?
```{r}
#Converting all datapoints to absolute values and taking the sum of each parameter
alignment %>% group_by(variable) %>% summarise(sum(sqrt(value^2)))
```


### 3.c. Are any of the realignment parameters significantly correlated with the fMRI model (same model as used in exercise 3)?
```{r}
#Transforming the data back from melted format and into a matrix
alignment <- read.delim("alignment.txt", header = F, sep = "")
alignment <- as.matrix(alignment)

#Making the correlation test between all variables
cor <- rcorr(alignment, fmrides, "spearman")

#Extracting the relevant values
cor <- data.frame("Rho_V1" = cor$r[1:6,7],
                 "p_value_V1" = cor$P[1:6,7],
                 "Significant_V1" = ifelse(cor$P[1:6,7]<0.05,"*"," "),
                 "Rho_V2" = cor$r[1:6,8],
                 "p_value_V2" = cor$P[1:6,8],
                 "Significant_V2" = ifelse(cor$P[1:6,8]<0.05,"*"," "))
cor
```
Realignment parameter 4 (pitch) is significantly correlated with the second model covariate, which most likely is due to chance. 

### Remove linear effects of time from the realignment parameters (hint: 1:400, fit a line and use residuals).
```{r}
#Making a time vector
time <- c(1:400)

#Making a linear model
m <- lm(alignment~time)

#Encoding the residuals to matrix
residuals <- m$residuals

```


### 3.d. Make a lineplot of the realignment parameters with time removed.
```{r}
#Converting to new format
melt_residuals <- melt(residuals)

#Making the lineplot
ggplot(melt_residuals, aes(Var1, value, colour = Var2))+
  geom_line()+
  labs(x = "Picture no.", y = "Value", title = "Realignment parameters with time removed")
```


### 3.e. Make a lineplot including only the first realignment parameter before and after removal.
```{r}
#Making a new data frame with the residuals and values of the first parameter. 
V1 <- data.frame("Time_corrected" = melt_residuals$value[melt_residuals$Var2 == "V1"],
                 "Not_time_corrected" = alignment[,1])

#Transforming the data frame into a new format
V1 <- melt(V1)

#Adding a variable for the x-axis
V1$Picture_no <- c(1:400)

#Plotting 
ggplot(V1, aes(Picture_no, value, colour = variable))+
  geom_line()+
  ggtitle("First realignment parameter")
```

### 3.f. Are the realignment parameters (corrected for effects of time) now correlated with the fMRI model?

```{r}
#Doing the correlation test like earlier
res_cor <- rcorr(residuals, fmrides, "spearman")

#Extracting the relevant correlation tests. 
res <- data.frame("Rho_V1" = res_cor$r[1:6,7],
                 "p_value_V1" = res_cor$P[1:6,7],
                 "Significant_V1" = ifelse(res_cor$P[1:6,7]<0.05,"*"," "),
                 "Rho_V2" = res_cor$r[1:6,8],
                 "p_value_V2" = res_cor$P[1:6,8],
                 "Significant_V2" = ifelse(res_cor$P[1:6,8]<0.05,"*"," "))

res
```
There are 3 significant relations between the second covariate and the parameters. They are probably due to change, but the first p value at 0.00011 seems very unlikely to get by chance, so maybe there is an actual correlation?

