---
title: "The_final_code"
author: "Anders Hjulmand, Morten Gade, Sigrid Nielsen and Gustav Helms"
date: "2/3/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Setup
library(pacman)
pacman::p_load("tidyverse")

#Loading the data
df <- read.csv("sleepstudy.csv", header = T)
df$Subject <- as.factor(df$Subject)
```

## Task 1
### 1.a: Get the data from one participant, e.g. using subset(). Make a linear regression for reaction time as a function of days of sleep deprivation, e.g. using lm(). Report the F-statistics.
```{r}
#Making a subset df
participan1 <- subset(df, df$Subject == 308)

#Making a scatterplot of subject 308
ggplot(participan1, aes(Days, Reaction))+
  geom_point()+
  geom_smooth(method = lm, alpha = 0.2)+
  ggtitle("Scatterplot of subject 308's reaction time by days of sleep deprivation")

#Linear regression model
summary(lm(participan1$Reaction~participan1$Days))
```
A linear regression analysis was used to test if the amount of days of sleep deprivation significantly predicted reaction time. The results of the regression indicated that the predictor explained 64.17% of the variance in the reaction time (Adjusted R2 =.6417, F(1,8)= 17.12, p<.01). It was found that a larger duration of sleep deprivation significantly predicted reaction time tendencies (β = 21.77, SE =5.26, t = 4.137, p<.01)


### 1.b: How many degrees of freedom does the relevant F-distribution have?
The relevant F-distribution have df=1 for predictors and df=8 for subjects. 

### 1.c: At which F-value does a regression with this distribution become statistically significant (p<0.05)?
```{r}
#Getting the F-score for the 95% significance cutoff using the quantile function qf()
cutoff<-qf(0.95,1,8)
'p=0.05 cutoff value with df(1,8)'
cutoff
```
A F-distributio with df(1,8) becomes statistically significant when F>5.318 at a significance level of 0.05. 

### 1.d: Make a plot of the F-distribution.
```{r}
#The F statistics for the effect of days on reaction time
model<-lm(participan1$Reaction~participan1$Days)
res=anova(model)
#Making a string of 100 numbers beteen 0 and 20 to use when finding the F distribution
nn<-seq(0,20,len=100)
#Getting the F distribution using df()
fdist<-df(nn,1,8)
plot(nn,fdist, type='l',xlab='F value - df(1,8)',ylab='Density')
#plotting a vertical line at the cutoff
lines(c(cutoff,cutoff),c(0,1),col='darkgreen')
#Add explanation for the line
text(cutoff+2,0.5,'p=0.05',col='darkgreen')
#draw F-value as point on the curve
points(res$`F value`[1],df(res$`F value`[1],1,8),col='red')
#add text for the f-value
text(res$`F value`[1]+2,0.1,'f=17,12',col='darkgreen')

```

## Task 2. For all participant in the experiment
### 2.a: Find the coefficients (slope and intercept) for the regression for reaction time as a function of days of sleep deprivation (a hint for the solution: use group_by() in tidyverse or this function here: https://stat.ethz.ch/R-manual/R-devel/library/nlme/html/lmList.html,hint2: pool=FALSE)
```{r}
all_subjects <- df %>% group_by(Subject) %>% summarise("Intercept" = lm(Reaction~Days)$coefficients[1],
                                                       "Slope" = lm(Reaction~Days)$coefficients[2])
all_subjects
```

### 2.b: Combine both scatter plot and regression line in the same figure. You may also include all participants in one plot.
```{r}
#Making the scatterplot
ggplot(df, aes(Days,Reaction, colour = Subject))+
  geom_point()+
  geom_smooth(method = lm, alpha = 0, size = 0.5)+
  ggtitle("Scatterplot and regression line for each subject")
```

### 2.c: Collect and report the inferential statistics for each participant in a table using t-statistics, including t-value, df and p-value.
### 2.d: How many individual participants display a statistically significant effect of sleep deprivation (p-values uncorrected for mulitple comparisons)?
```{r}
#Adding the inferential t-statistics to the df
all_subjects <- df %>% group_by(Subject) %>% summarise("Intercept" = lm(Reaction~Days)$coefficients[1],
                                                       "t-value_intercept" = summary(lm(Reaction~Days))$coefficients[1,3],
                                                       "p-value_intercept" = summary(lm(Reaction~Days))$coefficients[1,4],
                                                       "Slope" = lm(Reaction~Days)$coefficients[2],
                                                       "t-value_slope" = summary(lm(Reaction~Days))$coefficients[2,3],
                                                       "p-value_slope" = summary(lm(Reaction~Days))$coefficients[2,4],
                                                       "df" = lm(Reaction~Days)$df.residual)

#Adding a collumn with marked "*" if p<0.05.
all_subjects$significant <- ifelse(all_subjects$`p-value_slope`<0.05,"*"," ") 

all_subjects
```

## Task 3: Across participants: 
### 3.a: Use the slopes you found for each participant in exercise 2 as a new dataset. Test the hypothesis that the slopes are larger than zero against the null-hypothesis that the slopes are zero (i.e. no differences in response time exist as a function of time).

```{r}
#Testing for normality 
hist(all_subjects$Slope)
#By visual inspection the data approximates normal distribution even though the small sample size of n=18.
```

```{r}
#Making a 1-sample t-test to test if the slopes significantly differs from 0.       
m <- t.test(all_subjects$Slope)
m
```
### 3.b: Justify your use of test statistics.
T-test is used to test whether two means are statically significant different from each other. We used a 1-sample t-test of the slopes to test if they differed from 0, and in that way if tere is a significant of sleep deprivation on reaction time. 

### 3.c: Report inferential statistics.
Using an 1-sample t-test, we found that the mean gradient of the slopes significantly differed from 0, t(17) = 6.77, p<0.001, (M slope = 10.47, M = 0), which indicates that sleep deprication significantly modulates reaction time. 

### 3.d: Make a plot with the mean reaction time and standard errror bars for each day across participants and plot the averaged regression line in the same figure.

```{r}
#Making a plot
ggplot(df, aes(Days, Reaction))+
  stat_summary(fun.y = mean, geom = "point", colour = "black")+
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.3)+
  geom_smooth(method = lm, alpha = 0)+
  ggtitle("Mean reaction time pr. day across participants with se-bars")
```

